---
title: "AnalyzeData.DeshpandeM-UllalK.Rmd"
author: "Madhusudan Malhar Deshpande"
date: '2023-04-20'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Required Packages

This code checks if the required packages are already installed. If not, the packages are installed and then loaded

```{r load_packages}
# Package names
packages <- c("RMySQL", "ggplot2")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```


## Connect to Database

### Please put your database username and password before running this code

```{r DB_Connection_Setup, include=TRUE}

#1. Settings
db_user <- 'root'
db_password <- 'prakar@9599'
db_name <- 'pubmed'
db_host <- '127.0.0.1'
db_port <- 3306

#2. Read data from db
db_conn <- dbConnect(RMySQL::MySQL(), user=db_user, password = db_password, 
                     dbname = db_name, host = db_host, port = db_port)
```


## Analytical Query I

Top five journals with the most articles published in them for the time period

```{r}

sql <- "
  SELECT journal_sk, journal_title, sum(article_count) as articles_published
  FROM Jounrals_Fact
  GROUP BY journal_sk, journal_title
  ORDER BY articles_published DESC
  LIMIT 5
"

rs <- dbGetQuery(db_conn, sql)
head(rs)

ggplot(rs, aes(x=journal_title, y=articles_published)) + 
  geom_bar(stat = "identity") +
  labs(x="Journal", y="#Articles") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  geom_text(aes(label=articles_published), position=position_dodge(width=0.9), 
            vjust=-0.25) + 
  ylim(0,max(rs$articles_published)*1.1)

```


## Analytical Query II

Number of articles per journal per year broken down by quarter.

```{r}

sql <- "
  SELECT B.journal_sk, B.journal_title, A.quarter, A.year, B.article_count
  FROM (SELECT distinct quarter, year from Date_Dim) A 
  LEFT JOIN Jounrals_Fact B
  ON A.quarter = B.pubDate_quarter
  AND A.year = B.pubDate_year
"

rs <- dbGetQuery(db_conn, sql)
head(rs)
```


Report: 

The following 5 journals had the most publications in the given time frame. The frame ranges from 1975 to 1979. 

1. The Journal of pharmacy and pharmacology (1036)
2. Biochimica et biophysica acta (920)
3. The Journal of biological chemistry (604)
4. Annales de l'anesthesiologie francaise (542)
5. Biochemistry (375). 

They are represented in the bar graph above. 

The number of articles published for each journal in each of the quarter for the given time period is stored in a data frame above. Since this data is given for each journal, storing them in a dataframe is the best way to represent it. We can then filter for any particular journal to dig deeper whenever required.

